\documentclass[a5paper,11pt,english]{article}
\usepackage{.styles/basic}
\usepackage{.styles/envs}

%%%%%%% Title %%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{\textbf{Algebraic Geometry} - Exercise Sheet 8}
\author{Tor Gjone \& Paul}

%%%%%%% Definitions %%%%%%%%%%%%%%%%%%%%%
\input{defs}

\newSimpleHeaderEnvironment{exercise}{Exercise }

\DeclareMathOperator{\Fun}{Fun}
\DeclareMathOperator{\Spec}{Spec}

\DeclareMathOperator{\coker}{coker}

\renewcommand{\O}{\mathcal{O}}
\DeclareMathOperator{\QCoh}{QCoh}


\renewcommand{\E}{\mathcal{E}}
\renewcommand{\F}{\mathcal{F}}

\newcommand{\M}{\mathcal{M}}
\renewcommand{\N}{\mathcal{N}}

%%%%%%%% Content %%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\mmaketitle


\begin{exercise}[1]
\begin{enumerate}
\item
By definition $\QCoh(X)$ is a full subcategory of $\O_X$-Mod, since a morphism
in $\QCoh(X)$ is simply a morphism of $\O_X$-modules. So (assuming $\O_X$ is
abilian) we only need to check that the kernels and cokernels of morphisms of
quasi-coherent $\O_X$-modules are in $\QCoh(X)$.

Let $f : \M \to \N$, be a morphism of quasi-coherent $\O_X$-modules. 
Suppose $X = \Spec A$. Then, by Corollary 11.7 in the notes, 
$\ker f = \til{\ker (f(X))}$ and $\coker \til{\coker (f(X))}$, and are
thus quasi-coherent.

Now let $(U_i)_{i\in I}$ be a cover of $X$ by affines, such that $\M_{|U_i}
\cong \til {M_I}$ and $\N_{|U_i} \cong \til{N_i}$, for $\O_X(U_i)$-modules
$M_i$ and $N_i$.

Then, by the above argument, $\ker f_{|U_i} \cong \til{\ker f(U_i)}$, so $\ker
f$ is quasi-coherent. Similarly, $\coker f_{|U_i} \cong \til{\coker
f(U_i)}$, so also $\coker f$ is quasi-compact.


To show that $\QCoh(X)$ is closed under extensions, let 
\[ 0 \to \M \xto{f} \E \xto{g} \N \to 0 \]
be an exact sequence in $\O_X$-Mod, with $\M,\N \in \QCoh(X)$.

First suppose $X = \Spec A$, then again, by corollary 11.7,  
\[ 0 \to M \xto{f(X)} \E(X) \xto{g(X)} N \to 0 \]
is exact, with $N,M$ $A$-modules. As $\E$ is an $\O_X$-module, $\E(X)$ is also
an $A$-module. Applying the tildification again we get, by cor 11.7, that 
\[ 0 \to \til M \xto{\til{f(X)}} \til{\E(X)} \xto{\til{g(X)}} \til N \to 0 \]
is exact. Since $\M \cong \til M$, $\N \cong \til N$, $f = \til{f(X)}$ and $g =
\til{g(X)}$, we get that also $\E = \til {\E(X)}$.


If $(U_i)_{i\in I}$ is a cover of $X$, by affines such that $\M_{|U_i} \cong
\til {M_i}$ and $\N_{|U_i} \cong \til {N_i}$. Then, since exactness is a local
property (we could consider stalks),
\[ 0 \to \M_{|U_i} \xto{f_{|U_i}} \E_{|U_i} \xto{g_{|U_i}} \N_{|U_i} \to 0 \]
is also exact, and thus by the argument above $\E_{|U_i} \cong \til {\E(U_i)}$.


Finally it is immediate from the fact that the functor is inclusion and that it is
fully frightful, that it is exact.

\item
Since the kernel and cokernel of morphisms between finally generated modules are
also f.g., it is clear that the statement about abelian also hold for coherent
$\O_X$-modules. Similarly the category of f.g. modules is closed under
extensions, so also this statement holds for coherent $\O_X$-modules. 

%The exactness is also clear 
\end{enumerate}
\end{exercise}

\newpage

% \begin{exercise}[2]
% \begin{enumerate}
% \item
% Let $\gamma : \E \to \F$ be an isomorphism of vector bundles. 
% Let $\alpha' : \O^n_{U_0} \to \F_{|U_0}$ and $\beta' : \O^n_{U_1} \to
% \F_{|U_1}$. We have the diagram
% 
% \[ \begin{tikzcd}
% \O^n_{U_0\cap U_1} \arrow[r, "\alpha_|" anchor=south] \arrow[dr, "\alpha'_|"  anchor=north
% east] &
% \E_| \arrow[d, "\gamma_|"] &
% \O^n_{U_0 \cap U_1} \arrow[l, "\beta_|" anchor=south] \arrow[dl,"\beta'_|"] \\
% & \F_|
% \end{tikzcd} \]
% where $\cdot_| := \cdot_{|U_0\cap U_1}$. 
% The diagram does not commutes in general, but the left tirangle
% comumutes up to 
% 
% 
% 
% 
% 
% 
% \end{enumerate}
% \end{exercise}




\end{document}
